---
title: "Project Memo"
name: "Anna Bauer, Grant Cai, Alexis Navarra"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

install.packages("spotifyr")
library(spotifyr)

library(tidyverse)
library(knitr)
```

```{r}
install.packages('lubridate')
library(lubridate)
install.packages("httpuv")
library(httpuv)
```

## Overview of our dataset

We are using a data set called the Million Songs dataset as our training data. We are obtaining our test data through one of our group member's spotify accounts. We are able to access this data using an R package called spotifyr. The package contains many different functions that grant us access to data pertaining artists, tracks, albums, playlists, and more. This is a hyperlink to access this package: [Spotify Dataset](https://www.rcharlie.com/spotifyr/).
```{r, echo = F}
Sys.setenv(SPOTIFY_CLIENT_ID = 'da295936d8a2457b8eae51d982b99425')
Sys.setenv(SPOTIFY_CLIENT_SECRET = '9a802ea0578a4deeb0cdab988aae206a')

access_token <- get_spotify_access_token()

joy <- get_artist_audio_features('joy division')
head(joy)

joy %>% 
    arrange(-valence) %>% 
    select(track_name, valence) %>% 
    head(5) %>% 
    kable()
```


```{r}
get_my_top_artists_or_tracks(type = 'tracks', time_range = 'short_term', limit = 5) %>% 
    mutate(artist.name = map_chr(artists, function(x) x$name[1])) %>% 
    select(name, artist.name, album.name) %>% 
    kable()

```

```{r}
get_my_top_artists_or_tracks(type = 'artists', time_range = 'long_term', limit = 5) %>% 
    select(name, genres) %>% 
    rowwise %>% 
    mutate(genres = paste(genres, collapse = ', ')) %>% 
    ungroup %>% 
    kable()

```

```{r}
my_id <- 'abauer726'
my_plists <- get_user_playlists(my_id)

my_plists2 <- my_plists

tracks <- get_playlist_tracks(my_plists2$id) %>%
    filter(name %in% c('Calm Vibes',  'dance final'))
features <- get_track_audio_features(tracks)
```
## Overview of our research question

We are interested in predicting if someone will skip a song based on their music preference 

## Proposed project timeline and group work



## Any questions or concerns
