---
title: "Project Memo"
name: Anna Bauer, Grant Cai, Alexis Navarra
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
# install.packages("spotifyr")
# install.packages("httpuv")
# install.packages('lubridate')
```

```{r, echo=FALSE}
library(spotifyr)
library(tidyverse)
library(knitr)
library(lubridate)
library(httpuv)
```

## Overview of our dataset

We are finding our training dta from a subset of the Spotify Million Songs Dataset (the original dataset was far too large for our computers). [Training Data](https://corgis-edu.github.io/corgis/csv/music/)/. The data includes Spotify's chosen predictors such as artist familiarity, artist location, song id, song duration, etc. A preview of our data set is provided below. 
```{r}
data <- read.csv("https://raw.githubusercontent.com/abauer726/DoubleProjectGroup/main/131%20Project/Spotify%20Dataset/music.csv?token=GHSAT0AAAAAABQVVO3OPO66XXKMIJFCVAKWYPJ74BQ")


head(data)
dim(data)
```

This data set has 10,000 observations and 35 predictors.


We are obtaining our test data through one of our group member's spotify accounts. We are able to access this data using an R package called spotifyr. The package contains many different functions that grant us access to data pertaining artists, tracks, albums, playlists, and more. This is a hyperlink to access this package: [Spotify Dataset](https://www.rcharlie.com/spotifyr/).
```{r, echo = F}
Sys.setenv(SPOTIFY_CLIENT_ID = 'da295936d8a2457b8eae51d982b99425')
Sys.setenv(SPOTIFY_CLIENT_SECRET = '9a802ea0578a4deeb0cdab988aae206a')

access_token <- get_spotify_access_token()

```

In order to test that we are able to access the data from Anna's spotify account, we ran some code to see if we could retrieve her top 5 songs at the moment. The results are shown below.
```{r, echo = T}
get_my_top_artists_or_tracks(type = 'tracks', time_range = 'short_term', limit = 5) %>% 
    mutate(artist.name = map_chr(artists, function(x) x$name[1])) %>% 
    select(name, artist.name, album.name) %>% 
    kable()

```

```{r, echo = FALSE, results = 'hide'}
get_my_top_artists_or_tracks(type = 'artists', time_range = 'long_term', limit = 5) %>% 
    select(name, genres) %>% 
    rowwise %>% 
    mutate(genres = paste(genres, collapse = ', ')) %>% 
    ungroup %>% 
    kable()

```

```{r, echo = FALSE, results='hide'}
my_id <- 'abauer726'
my_plists <- get_user_playlists(my_id)

my_plists2 <- my_plists

tracks <- get_playlist_tracks(my_plists2$id) %>%
    filter(name %in% c('Calm Vibes',  'dance final'))
features <- get_track_audio_features(tracks)
```

We are working with mostly quantitative data, which leads us to believe that regression may be our most effective type of modeling. 

Missing data: All of the song titles are missing, but the good news is that there are specific song id's that make each observation unique. We will conduct a test in the near future to find out how many values are missing in each column. 

We only want to work with the significant predictors in the data set, which we will find in our initial analysis. 

## Overview of our research question

Our research question is: are we able to predict whether or not someone will want to skip a song based on their music taste and certain qualities of the song?

Our response variable is skip/no skip; will a random person skip the song, or will they listen through the song?


We believe that our research question will be best answered with a regression approach. We believe that the predictors beats per minute, artist familiarity, tempo, loudness, and duration will be especially useful in answering our research question.



## Proposed project timeline and group work

For a general timeline, we are going to follow closely along to the one that Professor Coburn attached to the syllabus. Here is a run-down of our timeline:

Week 1: Find team + topic
Week 2: Find data
Week 3: Load data, tidy data, come up with scope of the project
Week 4: Visualization 
Week 5: Initial analysis
Week 6: Initial analysis
Week 7: Run models 
Week 8: Run models
Week 9: Write up results
Week 10: Write up results

Our group is going to be meeting weekly on Tuesdays at 12pm, where we will take a group look at what we need to accomplish for the week and delegate weekly tasks.

## Any questions or concerns

We anticipate that this project is going to be difficult, as we are all generally new to coding + machine learning, so we will bring up questions throughout the quarter as needed. 