---
title: "Project Memo"
name: "Anna Bauer, Grant Cai, Alexis Navarra"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

install.packages("spotifyr")
library(spotifyr)
```

## Overview of our dataset

We have our data installed in an R package called spotifyr. The package contains many different functions that grant us access to data pertaining artists, tracks, albums, playlists, and more. This is a hyperlink to our data: [Spotify Dataset](https://www.rcharlie.com/spotifyr/).
```{r, echo = F}
Sys.setenv(SPOTIFY_CLIENT_ID = 'da295936d8a2457b8eae51d982b99425')
Sys.setenv(SPOTIFY_CLIENT_SECRET = '9a802ea0578a4deeb0cdab988aae206a')

access_token <- get_spotify_access_token()

joy <- get_artist_audio_features('joy division')
head(joy)

joy %>% 
    arrange(-valence) %>% 
    select(track_name, valence) %>% 
    head(5) %>% 
    kable()

get_my_recently_played(limit = 5) %>% 
    mutate(artist.name = map_chr(track.artists, function(x) x$name[1]),
           played_at = as_datetime(played_at)) %>% 
    select(track.name, artist.name, track.album.name, played_at) %>% 
    kable()
```

## Overview of our research question

We are interested in predicting if someone will skip a song based on their music preference 

## Proposed project timeline and group work



## Any questions or concerns
