---
title: "Project Memo"
name: Anna Bauer, Grant Cai, Alexis Navarra
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
# install.packages("spotifyr")
# install.packages("httpuv")
# install.packages('lubridate')
```

```{r, echo=FALSE}
library(spotifyr)
library(tidyverse)
library(knitr)
library(lubridate)
library(httpuv)
```

## Overview of our dataset

We are using a subset of a data set called the Million Songs data set as our training data. A preview of our data set is provided below.
```{r}
data <- read.csv("https://raw.githubusercontent.com/abauer726/DoubleProjectGroup/main/131%20Project/Spotify%20Dataset/music.csv?token=GHSAT0AAAAAABQVVO3OPO66XXKMIJFCVAKWYPJ74BQ")



head(data)
dim(data)
```



This data set has 10,000 observations and 35 predictors.


We are finding our training dta from a subset of the Spotify Million Songs Dataset (the original dataset was far too large for our computers). [Training Data](https://corgis-edu.github.io/corgis/csv/music/)

We are obtaining our test data through one of our group member's spotify accounts. We are able to access this data using an R package called spotifyr. The package contains many different functions that grant us access to data pertaining artists, tracks, albums, playlists, and more. This is a hyperlink to access this package: [Spotify Dataset](https://www.rcharlie.com/spotifyr/).
```{r, echo = F}
Sys.setenv(SPOTIFY_CLIENT_ID = 'da295936d8a2457b8eae51d982b99425')
Sys.setenv(SPOTIFY_CLIENT_SECRET = '9a802ea0578a4deeb0cdab988aae206a')

access_token <- get_spotify_access_token()

```

In order to test that we are able to access the data from Anna's spotify account, we ran some code to see if we could retrieve her top 5 songs at the moment. The results are shown below.
```{r, echo = T}
get_my_top_artists_or_tracks(type = 'tracks', time_range = 'short_term', limit = 5) %>% 
    mutate(artist.name = map_chr(artists, function(x) x$name[1])) %>% 
    select(name, artist.name, album.name) %>% 
    kable()

```

```{r, echo = FALSE, results = 'hide'}
get_my_top_artists_or_tracks(type = 'artists', time_range = 'long_term', limit = 5) %>% 
    select(name, genres) %>% 
    rowwise %>% 
    mutate(genres = paste(genres, collapse = ', ')) %>% 
    ungroup %>% 
    kable()

```

```{r, echo = FALSE, results='hide'}
my_id <- 'abauer726'
my_plists <- get_user_playlists(my_id)

my_plists2 <- my_plists

tracks <- get_playlist_tracks(my_plists2$id) %>%
    filter(name %in% c('Calm Vibes',  'dance final'))
features <- get_track_audio_features(tracks)
```

We only want to work with a portion of the predictors in the data set: ___. 
In our data set, there is ___ missing data

## Overview of our research question

Our research question is: are we able to predict whether or not someone will want to skip a song based on their music taste and certain qualities of the song?
Our response variable is ___.
We believe that our research question will be best answered with a classification approach. We believe that the predictors ___ will be especially useful in answering our research question.

## Proposed project timeline and group work

For a general timeline, we are going to follow closely along to the one that Professor Coburn attached to the syllabus. Here is a run-down of our timeline:

Week 1: Find team + topic
Week 2: Find data
Week 3: Load data, tidy data, come up with scope of the project
Week 4: Visualization 
Week 5: Initial analysis
Week 6: Initial analysis
Week 7: Run models 
Week 8: Run models
Week 9: Write up results
Week 10: Write up results

Our group is going to be meeting weekly on Tuesdays at 12pm, where we will take a group look at what we need to accomplish for the week and delegate weekly tasks.

## Any questions or concerns

We anticipate that this project is going to be difficult, as we are all generally new to coding + machine learning, so we will bring up questions throughout the quarter as needed. 